---
title: "Capstone"
author: "Michael Taffe"
date: "11/15/2021"
output:
  pdf_document: default
  word_document: default
---

# Reading Data
```{r}
cs.data = read.table('./Data4.tsv', sep = '\t', header = TRUE) 
head(cs.data)

library(tidyverse)
```
# interpreting variables
```{r}
summary(cs.data)
```
\ All x variables from summary look to be skewed. We will plot histograms to further analyse

```{r}
par(mfrow=c(1,3))
hist(cs.data$Tempchange)
hist(cs.data$Population)
hist(cs.data$Area)
hist(cs.data$Density)
hist(cs.data$Coast)
hist(cs.data$Latitude)
hist(cs.data$Altitude)
```
\ Population, Area, Density, Coast, and altitude are all skewed right
### Box-cox (To conduct this test we will need to take the abs of Temp change as Boxcox needs positive response variable)
```{r}
library(MASS)
```
```{r}
cs.data$abstemp = abs(cs.data$Tempchange)
```

```{r}
Popslr = lm(abstemp ~ Population, data=cs.data)
Areaslr = lm(abstemp~ Area, data=cs.data)
Densslr = lm(abstemp ~ Density, data=cs.data)
Altslr = lm(abstemp ~ Altitude, data=cs.data)
```

```{r}
boxcox(Popslr)
```
```{r}
boxcox(Areaslr)
```
```{r}
boxcox(Densslr)
```
```{r}
boxcox(Altslr)
```
### Linearizing without boxcox 
```{r}
Linpop = log(cs.data$Population)
Linarea = log(cs.data$Area)
Lindense = log(cs.data$Density)
Linalt = log(cs.data$Altitude)
```
```{r}
par(mfrow=c(1,2))
hist(Linpop)
hist(Linarea)
hist(Lindense)
hist(Linalt)
```
\ These variables follow a much more normal distribution and will be used in building the models.
### Testing for outliers
forgot to do this oops

# Geographic Model
## Building Models
```{r}
Geomlr = lm(Tempchange ~ Latitude + Altitude + Coast, data = cs.data)
```
```{r}
summary(Geomlr)
```
```{r}
anova(Geomlr)
```

### Best Subset
```{r}
library(leaps)
```
```{r}
temp = regsubsets(Tempchange ~ Latitude + Altitude + Coast, data=cs.data)
bestsub = summary(temp)
bestsub
```
## Plotting Model

## Testing Assumptions 
```{r}
library(car)
```
```{r}
avPlots(Geomlr)
```
```{r}
vif(Geomlr)
```
```{r}
influenceIndexPlot(Geomlr)
```
```{r}
rstudent(Geomlr)
```
```{r}
dffits(Geomlr)
```
```{r}
dfbetas(Geomlr)
```
```{r}
n <- 100 # sample size
p <- 4 # 4 parameters (1 intercept, 3 slopes)
par(mfrow=c(2,2))
plot(rstudent(Geomlr), ylab="studentized deleted residual")
abline(qt(1-0.1/(2*n), df=n - p - 1), 0, lty=2)
abline(qt(0.1/(2*n), df=n - p - 1), 0, lty=2)
plot(dffits(Geomlr), ylab = "dffits")
abline(1, 0, lty=2)
abline(1, 0, lty=2)
plot(cooks.distance(Geomlr), ylab="Cook's Distance")
abline(qf(0.5, df1=5, df2=n-p), 0, lty=2)
plot(hatvalues(Geomlr), ylab="leverage")
abline(2*p/n, 0, lty=2)
```
```{r}
par(mfrow=c(2,3))
for (i in 1:4) {
plot(dfbetas(Geomlr)[, i], ylab=paste("Dfbeta_", i-1))
abline(1, 0, lty=2)
abline(-1, 0, lty=2)
}
```
```{r}
library(lmtest)
```
```{r}
dwtest(Geomlr)
```
```{r}
dwtest(Geomlr, alternative = "two.sided")
```
\ P value indicates that there is not 1st order auto correlation
```{r}
qqnorm(Geomlr$residuals)
qqline(Geomlr$residuals)
```

# City Design Model
## Building Models
```{r}
Lintrue = log(cs.data$TrueSize)
```

```{r}
Desmlr = lm(Tempchange ~ Population + Density + Area + Lintrue, data = cs.data)
```
```{r}
Desmlr
```

```{r}
summary(Desmlr)
```
```{r}
anova(Desmlr)
```

### Best Subset
```{r}
temp2 = regsubsets(Tempchange ~ Population + Area + Density + TrueSize, data=cs.data)
bestsub2 = summary(temp2)
bestsub2
```
## Plotting Model

## Testing Assumption
```{r}
avPlots(Desmlr)
```
```{r}
vif(Desmlr)
```
```{r}
influenceIndexPlot(Desmlr)
```
```{r}
rstudent(Desmlr)
```
```{r}
dffits(Desmlr)
```
```{r}
dfbetas(Desmlr)
```
```{r}
n <- 100 # sample size
p <- 5 # 4 parameters (1 intercept, 3 slopes)
par(mfrow=c(2,2))
plot(rstudent(Geomlr), ylab="studentized deleted residual")
abline(qt(1-0.1/(2*n), df=n - p - 1), 0, lty=2)
abline(qt(0.1/(2*n), df=n - p - 1), 0, lty=2)
plot(dffits(Desmlr), ylab = "dffits")
abline(1, 0, lty=2)
abline(1, 0, lty=2)
plot(cooks.distance(Desmlr), ylab="Cook's Distance")
abline(qf(0.5, df1=5, df2=n-p), 0, lty=2)
plot(hatvalues(Desmlr), ylab="leverage")
abline(2*p/n, 0, lty=2)
```
\ Issues in the leverage
```{r}
par(mfrow=c(2,3))
for (i in 1:4) {
plot(dfbetas(Desmlr)[, i], ylab=paste("Dfbeta_", i-1))
abline(1, 0, lty=2)
abline(-1, 0, lty=2)
}
```
```{r}
dwtest(Desmlr)
```
```{r}
dwtest(Geomlr, alternative = "two.sided")
```
```{r}
qqnorm(Desmlr$residuals)
qqline(Desmlr$residuals)
```
# Combined Model
## Building Model
```{r}
cmmlr <- lm(Tempchange ~ Population + Area + Density + Coast + Latitude + Altitude + Lintrue, data = cs.data)
cmmlr
```
```{r}
summary(cmmlr)
```
```{r}
anova(cmmlr)
```
```{r}
avPlots(cmmlr)
```
```{r}
vif(cmmlr)
```
```{r}
influenceIndexPlot(cmmlr)
```
```{r}
rstudent(cmmlr)
```
```{r}
dffits(cmmlr)
```
```{r}
dfbetas(cmmlr)
```
```{r}
n <- 100 # sample size
p <- 8 # 8 parameters (1 intercept, 7 slopes)
par(mfrow=c(2,2))
plot(rstudent(cmmlr), ylab="studentized deleted residual")
abline(qt(1-0.1/(2*n), df=n - p - 1), 0, lty=2)
abline(qt(0.1/(2*n), df=n - p - 1), 0, lty=2)
plot(dffits(cmmlr), ylab = "dffits")
abline(1, 0, lty=2)
abline(1, 0, lty=2)
plot(cooks.distance(cmmlr), ylab="Cook's Distance")
abline(qf(0.5, df1=5, df2=n-p), 0, lty=2)
plot(hatvalues(cmmlr), ylab="leverage")
abline(2*p/n, 0, lty=2)
```

--------------------------------------------------------------------------------------------------

# Linearized Models

```{r}
Popslr_lnrz <- lm(abstemp ~ log(Population), data = cs.data)
Areaslr_lnrz <- lm(abstemp~ log(Area), data=cs.data)
Densslr_lnrz <- lm(abstemp ~ log(Density), data=cs.data)

cs.data$Altitude[84] <- 1
cs.data$Altitude[88] <- 1

Altslr_lnrz <- lm(abstemp ~ log(Altitude), data=cs.data)
```

```{r}
Desmlr_lnrz <- lm(Tempchange ~ log(Population) + log(Area) + log(Density) + log(TrueSize), data = cs.data)

summary(Desmlr_lnrz)
anova(Desmlr_lnrz)
```

```{r}
#temp2 = regsubsets(Tempchange ~ Population + Area + Density + Truesize, data=cs.data)
#bestsub2 = summary(temp2)
#bestsub2

temp2_lnrz <- regsubsets(Tempchange ~ log(Population) + log(Area) + log(Density) + log(TrueSize), data=cs.data)
bestsub2 <- summary(temp2_lnrz)
bestsub2
```

```{r}
summary(Desmlr)
```
```{r}
anova(Desmlr)
```

### Best Subset
```{r}
temp2 = regsubsets(Tempchange ~ Population + Area + Density + TrueSize, data=cs.data)
bestsub2 = summary(temp2)
bestsub2
```

## Plotting Model

## Testing Assumption
```{r}
avPlots(Desmlr_lnrz)
```
```{r}
influenceIndexPlot(Desmlr_lnrz)
```
```{r}
rstudent(Desmlr_lnrz)
```
```{r}
dffits(Desmlr_lnrz)
```
```{r}
dfbetas(Desmlr_lnrz)
```

```{r}
n <- 100 # sample size
p <- 5 # 4 parameters (1 intercept, 3 slopes)
par(mfrow=c(2,2))
plot(dffits(Desmlr_lnrz), ylab = "dffits")
abline(1, 0, lty=2)
abline(1, 0, lty=2)
plot(cooks.distance(Desmlr_lnrz), ylab="Cook's Distance")
abline(qf(0.5, df1=5, df2=n-p), 0, lty=2)
plot(hatvalues(Desmlr_lnrz), ylab="leverage")
abline(2*p/n, 0, lty=2)
```
\ Issues in the leverage
```{r}
par(mfrow=c(2,3))
for (i in 1:3) {
plot(dfbetas(Desmlr_lnrz)[, i], ylab=paste("Dfbeta_", i-1))
abline(1, 0, lty=2)
abline(-1, 0, lty=2)
}
```
```{r}
dwtest(Desmlr_lnrz)
```
```{r}
qqnorm(Desmlr_lnrz$residuals)
qqline(Desmlr_lnrz$residuals)
```

```{r}
cmmlr_lnrz <- lm(Tempchange ~ log(Population) + log(Area) + log(Density) + Coast + Latitude + Altitude + Lintrue, data = cs.data)
cmmlr
```
```{r}
summary(cmmlr_lnrz)
```
```{r}
anova(cmmlr_lnrz)
```
```{r}
avPlots(cmmlr_lnrz)
```
```{r}
influenceIndexPlot(cmmlr_lnrz)
```
```{r}
rstudent(cmmlr_lnrz)
```
```{r}
dffits(cmmlr_lnrz)
```
```{r}
dfbetas(cmmlr_lnrz)
```
```{r}
n <- 100 # sample size
p <- 8 # 8 parameters (1 intercept, 7 slopes)
par(mfrow=c(2,2))
plot(rstudent(cmmlr_lnrz), ylab="studentized deleted residual")
abline(qt(1-0.1/(2*n), df=n - p - 1), 0, lty=2)
abline(qt(0.1/(2*n), df=n - p - 1), 0, lty=2)
plot(dffits(cmmlr_lnrz), ylab = "dffits")
abline(1, 0, lty=2)
abline(1, 0, lty=2)
plot(cooks.distance(cmmlr_lnrz), ylab="Cook's Distance")
abline(qf(0.5, df1=5, df2=n-p), 0, lty=2)
plot(hatvalues(cmmlr_lnrz), ylab="leverage")
abline(2*p/n, 0, lty=2)
```





```{r}
summary(Geomlr)
```

As R Square adjusted tells us that this model only explains 9.53% of the variables these number is too low,
even though the p-value tells us that the overall model is statistically significant. Only the variable latitude and altitude has a significant t-value. It appears that coast isnt functioning as welll as the other variables.


```{r}
summary(Desmlr)
```


Althought Population and Lintrue are significant we will have to reject these model
as P-value associated to F-statistic isnt significant enough. We failed to reject reject the null hypothesis
that says that the overall model is not significant.


```{r}
summary(cmmlr)
```

These model appears to be the best,  we reject the null hypothesis associated
to F-stat as P-value is lower than alpha. Second, Adjusted-R square is higher than in the other two models
thus these model explaining 15.22% of temperature change. Third, we obtain a very significant variable
wich is latitude, and others highly significant such as Population, altitude and Lintrue.






```{r}
summary(Desmlr_lnrz)
```


Althought Population and Lintrue are significant we will have to reject these model
as P-value associated to F-statistic isnt significant enough. We failed to reject reject the null hypothesis
that says that the overall model is not significant. 



```{r}
summary(cmmlr_lnrz)
```


These model appears to be the second best, we reject the null hypothesis associated
to F-stat as P-value is lower than alpha 0.05. Second, Adjusted-R square is 0.117 meaning this model explains 11.70% of data.  Variable latitude was highly signficant and variables altitude and log(area) seems to help the model too.




```{r}
summary(csmlr3)
```

These model appears to be the third best, we reject the null hypothesis associated
to F-stat as P-value is lower than alpha 0.05. Second, Adjusted-R square is lower than the two best models explaining only 9.53% of the variables. Third, we loss a lot of information as these model does not contemplate important variables such as population and Lintrue, and also the variable latitude looses significance.

Calculating AIC of top 3 models (cmmlr,cs.mlr1, and csmlr3)

```{r}
library(tidyverse)
library(broom)



```


```{r}
library(tidyverse)
library(broom)
library(modelr)

AIC(cmmlr)
BIC(cmmlr)
```
```{r}
AIC(cmmlr_lnrz)
BIC(cmmlr_lnrz)
```

```{r}
AIC(Geomlr)
BIC(Geomlr)
```